---
title: "Untitled"
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

---
title: "Fitness Tracker Trend Analysis"
author: "Adelia Ramadhani"
date: "1/11/2022"
output: html_document
---

### **Overview**

[Bellabeat](https://bellabeat.com/?) is a women-centric tech and
wellness company which develops wearables and accompanying products that
monitor biometric and lifestyle data to help women better understand how
their bodies work, and as a result, make healthier lifestyle choices.

Together with the Bellabeat app, users are able to gain insights with
health data related to their activity, sleep, stress, fitness, heart
rate, reproductive health and mindfulness habits.

The aim of this project is to analyze how non-Bellabeat consumers use
their smart fitness devices. With this information, we are to provide
high-level recommendations for how these insights can inform Bellabeat's
marketing strategy.

This analysis consists of 6 phase: **Ask, Prepare, Process, Analyze,
Share** and **Act**.

### **Ask Phase**

To begin the analysis, we need to ask questions and problem-solve. In
this Ask phase, we need to identify the business task, the questions
that needs to be answered, and the stakeholders. These aspects will
guide us through the entire process of the data analysis.

#### **1. Business Task**

-   Analyze Fitbit data to gain insight into how consumers use
    non-Bellabeat smart devices.

-   Select one Bellabeat product and apply these insights into a
    presentation. *(In this case study, we choose **Leaf**).*

-   Provide high-level recommendations for Bellabeat's marketing
    strategy.

#### **2. Questions for The Analysis**

-   What are some trends in smart device usage?

-   How could these trends apply to Bellabeat customers?

-   How could these trends help influence Bellabeat's marketing
    strategy?

#### **3. Stakeholders**

-   Urška Sršen: Bellabeat's cofounder and Chief Creative Officer.

-   Sando Mur: Mathematician and Bellabeat's cofounder.

-   Bellabeat Marketing Analytics team.

### **Prepare Phase**

This phase will help us to make use the right data for our objectives
and how to organize and protect the data so that we can process them in
the next phase.

#### **1. Data Sources**

The data source used for our case study is FitBit Fitness Tracker Data.
This dataset is stored in Kaggle and was made available through Mobius.

-   Source Used: [FitBit Fitness Tracker
    Data](https://www.kaggle.com/arashnic/fitbit) (CC0: Public Domain
    Kaggle dataset made available through Mobius,
    04.12.2016-05.12.2016).

-   [Original Source:](https://zenodo.org/record/53894#.Ybzqy9nMK8X)
    Furberg, R., Brinton, J., Keating, M., & Ortiz, A. (2016).
    Crowd-sourced Fitbit datasets 03.12.2016-05.12.2016 Zenodo.

#### **2. Accessibility and Privacy of the Data**

Verifying the metadata of our dataset we can confirm it is open-source.
The owner has dedicated the work to the public domain by waiving all of
their rights to the work worldwide under copyright law, including all
related and neighboring rights, to the extent allowed by law. We can
copy, modify, distribute and perform the work, even for commercial
purposes, all without asking permission.

#### **3. Other Information about the Dataset**

-   These datasets were generated by respondents to a distributed survey
    via Amazon Mechanical Turk between 03.12.2016-05.12.2016 **(31
    days).**

-   **Thirty** eligible Fitbit users consented to the submission of
    personal tracker data, including minute-level output for physical
    activity, heart rate, and sleep monitoring.

-   Variation between output represents use of different types of Fitbit
    trackers and individual tracking behaviors / preferences.

#### **4. Data Organization and Verification**

Available to us are 18 CSV documents. Each document represents different
quantitative data tracked by Fitbit. The data is considered long since
each row is one time point per subject, so each subject will have data
in multiple rows.Every user has a unique ID and different rows since
data is tracked by day and time.

Due to the small size of sample I sorted and filtered tables creating
Pivot Tables in Excel. I was able to verify attributes and observations
of each table and relations between tables. Counted sample size (users)
of each table and verified time length of analysis - 31 days.

|          **Table Name**          |      **Type**       |                                                                          **Description**                                                                          |
|:--------------------------------:|:-------------------:|:-----------------------------------------------------------------------------------------------------------------------------------------------------------------:|
|      `dailyActivity_merged`      | Microsoft Excel CSV |                                  Daily Activity over 31 days of 33 users. Tracking daily: Steps, Distance, Intensities, Calories                                  |
|      `dailyCalories_merged`      | Microsoft Excel CSV |                                                              Daily Calories over 31 days of 33 users                                                              |
|    `dailyIntensities_merged`     | Microsoft Excel CSV | Daily Intensity over 31 days of 33 users. Measured in Minutes and Distance, dividing groups in 4 categories: Sedentary, Lightly Active, Fairly Active,Very Active |
|       `dailySteps_merged`        | Microsoft Excel CSV |                                                               Daily Steps over 31 days of 33 users                                                                |
|    `heartrate_seconds_merged`    | Microsoft Excel CSV |                                                        Exact day and time heartrate logs for just 7 users                                                         |
|     `hourlyCalories_merged`      | Microsoft Excel CSV |                                                          Hourly Calories burned over 31 days of 33 users                                                          |
|    `hourlyIntensities_merged`    | Microsoft Excel CSV |                                                    Hourly total and average intensity over 31 days of 33 users                                                    |
|       `hourlySteps_merged`       | Microsoft Excel CSV |                                                               Hourly Steps over 31 days of 33 users                                                               |
|  `minuteCaloriesNarrow_merged`   | Microsoft Excel CSV |                                        Calories burned every minute over 31 days of 33 users (Every minute in single row)                                         |
|   `minuteCaloriesWide_merged`    | Microsoft Excel CSV |                                       Calories burned every minute over 31 days of 33 users (Every minute in single column)                                       |
| `minuteIntensitiesNarrow_merged` | Microsoft Excel CSV |                                         Intensity counted by minute over 31 days of 33 users (Every minute in single row)                                         |
|  `minuteIntensitiesWide_merged`  | Microsoft Excel CSV |                                       Intensity counted by minute over 31 days of 33 users (Every minute in single column)                                        |
|    `minuteMETsNarrow_merged`     | Microsoft Excel CSV |                     Ratio of the energy you are using in a physical activity compared to the energy you would use at rest. Counted in minutes                     |
|       `minuteSleep_merged`       | Microsoft Excel CSV |                                             Log Sleep by Minute for 24 users over 31 days. Value column not specified                                             |
|    `minuteStepsNarrow_merged`    | Microsoft Excel CSV |                                         Steps tracked every minute over 31 days of 33 users (Every minute in single row)                                          |
|     `minuteStepsWide_merged`     | Microsoft Excel CSV |                                        Steps tracked every minute over 31 days of 33 users (Every minute in single column)                                        |
|        `sleepDay_merged`         | Microsoft Excel CSV |                                    Daily sleep logs, tracked by: Total count of sleeps a day, Total minutes, Total Time in Bed                                    |
|      `weightLogInfo_merged`      | Microsoft Excel CSV |            Weight track by day in Kg and Pounds over 30 days. Calculation of BMI.5 users report weight manually 3 users not.In total there are 8 users            |

#### **5. Data Limitations**

-   **No Metadata Provided:** Information such as location, lifestyle,
    weather, temperature, humidity etc. would provide a deeper context
    to the data obtained.

-   **Missing Demographics:** Key demographics data such as gender, age,
    were not identified. This is a crucial missing information sine
    Bellabeat creates women-centric products. Insights obtained may not
    reflect the differences in physiology and activity patterns between
    different demographic groups. However, we understand such
    information is under a strict privacy policy.

-   **Small Sample Size:** Thirty users is not an ideal sample size
    where multiple independent variables are involved. Especially when
    health and lifestyle data is varied across different facets of
    society. Insights gained may not apply to all.

-   **Data Collection Period:** 31 days of data is limited in providing
    high-level recommendations. Seasonal trends impacts heavily on user
    activity and lifestyle choices. E.g. User's exercise habits differ
    between summer and winter. Moreover, the data were collected 6 years
    ago, which is no longer current.

### **Process Phase**

In this phase, we will check, clean and transform the data, as well as
verify and report our cleaning results. Data processing, analysis and
visualization will all be done in R Programming with R Studio.

#### **1. Install Packages & Load Libraries Needed for Cleaning**

The libraries that we need to install are :

-   Tidyverse

-   Skimr

-   Tidyr

-   Lubridate

-   Janitor

-   Dplyr

-   Ggplot2

-   Ggpubr

-   Ggrepel

```{r install packages and load libraries}
options(repos = list(CRAN="http://cran.rstudio.com/"))
install.packages("tidyverse")
install.packages("skimr")
install.packages("tidyr")
install.packages("lubridate")
install.packages("janitor")
install.packages("dplyr")
install.packages("ggplot2")
install.packages("ggpubr")
install.packages("ggrepel")
install.packages("scales")
install.packages("waffle")
install.packages("plotly")
install.packages("viridis")

library("tidyverse")
library("skimr")
library("tidyr")
library("lubridate")
library("janitor")
library("dplyr")
library("ggplot2")
library("ggpubr")
library("ggrepel")
library("scales")
library("waffle")
library("plotly")
library("viridis")
```

#### **2. Import Datasets**

Knowing the datasets we have, we will upload the datasets that will help
us answer our business task. On our analysis we will focus on the
following datasets:

-   Daily_activity

-   Daily_sleep

-   Hourly_steps

-   Hourly_intensities

-   Hourly_calories

Due to the the small sample we won't consider for this analysis Weight
(8 Users) and heart rate (7 users)

```{r}
daily_activity <- read.csv("dailyActivity_merged.csv")
daily_sleep <- read.csv("sleepDay_merged.csv")
hourly_steps <- read.csv("hourlySteps_merged.csv")
hourly_intensities <- read.csv("hourlyIntensities_merged.csv")
hourly_calories <- read.csv("hourlyCalories_merged.csv")
```

#### **3. Preview The Datasets**

This step will give us the preview about the name of each column and the
type of each data frame so that we can know what kind of data that we
have in a brief.

```{r}
head(daily_activity)
str(daily_activity)

head(daily_sleep)
str(daily_sleep)

head(hourly_steps)
str(hourly_steps)

head(hourly_intensities)
str(hourly_intensities)

head(hourly_calories)
str(hourly_calories)
```

#### **4. Clean The Name of The Columns in Each Table**

The column names need to be formatted consistently to make analysis
easier.

```{r}
daily_activity <- clean_names(daily_activity)
daily_sleep <- clean_names(daily_sleep)
hourly_steps <- clean_names(hourly_steps)
hourly_intensities <- clean_names(hourly_intensities)
hourly_calories <- clean_names(hourly_calories)
```

#### **5. Remove Duplicates and N/A**

Data duplicates and N/A values can interfere with the analysis, so they
must be removed in order to make the data usable.

```{r}
daily_activity <- daily_activity %>% 
  distinct() %>% 
  drop_na()

daily_sleep <- daily_sleep %>% 
  distinct() %>% 
  drop_na()

hourly_steps <- hourly_steps %>% 
  distinct() %>% 
  drop_na()

hourly_intensities <- hourly_intensities %>% 
  distinct() %>% 
  drop_na()

hourly_calories <- hourly_calories %>% 
  distinct() %>% 
  drop_na()
```

After the entire process, we should check if the duplicates were

#### **6. Change The Faulty Data Types**

After knowing each data type of the data frame, we should have known if
there was any incorrect data type. In this step, we need to correct
them.

```{r}
daily_activity <- daily_activity %>%
  rename(date = activity_date) %>%
  mutate(date = as_date(date, format = "%m/%d/%Y"))
str(daily_activity)

daily_sleep <- daily_sleep %>%
  rename(date = sleep_day) %>% 
  mutate(date = as_date(date, format = "%m/%d/%Y"))
str(daily_sleep)

hourly_calories <- hourly_calories %>% 
  rename(date_time = activity_hour) %>% 
  mutate(date_time = as.POSIXct(date_time, format = "%m/%d/%Y %I:%M:%S %p", tz = Sys.timezone()))
str(hourly_calories)

hourly_intensities <- hourly_intensities %>% 
  rename(date_time = activity_hour) %>% 
  mutate(date_time = as.POSIXct(date_time, format = "%m/%d/%Y %I:%M:%S %p", tz = Sys.timezone()))
str(hourly_intensities)

hourly_steps <- hourly_steps %>% 
  rename(date_time = activity_hour) %>% 
  mutate(date_time = as.POSIXct(date_time, format = "%m/%d/%Y %I:%M:%S %p", tz = Sys.timezone()))
str(hourly_steps)
```

#### **7. Merge The Datasets**

Merging the datasets is needed to see if there is any correlation
between each data. In this step, we merge the datasets based on the time
it gets measured (daily or hourly).

```{r}
daily_activity_sleep <- merge(daily_activity, daily_sleep, by = c("id", "date"))
glimpse (daily_activity_sleep)

hourly_cal_intensities_steps <- hourly_calories %>% 
  left_join(hourly_intensities, by = c("id", "date_time")) %>% 
  left_join(hourly_steps, by = c("id", "date_time"))
glimpse(hourly_cal_intensities_steps)
```

#### **8. Check Min (Date), Max (Date), and N/A Values**

```{r}
min(daily_activity_sleep$date)
max(daily_activity_sleep$date)
n_unique(daily_activity_sleep$id)

min(hourly_cal_intensities_steps$date_time)
max(hourly_cal_intensities_steps$date_time)
n_unique(hourly_cal_intensities_steps$id)
```

### **Analysis and Share Phase**

We will analyze Fitbit user trends in order to come up with a better
marketing strategy for BellaBeat.

#### **1. Usage Group Distribution**

The purpose of this experiment is to determine how often participants
wear their watches. With `daily_activity_sleep`, we will determine the
type of users based on the daily `total_steps` they have taken. We will
assign the following designations:

-   Sedentary: \<5000 steps a day.

-   Lightly Active: 5000 - 7499 steps a day.

-   Fairly Active: 7500 - 9999 steps a day.

-   Very active: >10000 steps a day.

In making these classifications, we take into account the following
article: <https://www.10000steps.org.au/articles/counting-steps/>

We can gain a better understanding of each **Usage Group** by further
breaking down the analysis in this way.

-   Calculate the daily steps average for every user:

```{r}
daily_average <- daily_activity_sleep %>%
  group_by(id) %>%
  summarise (mean_daily_steps = mean(total_steps), mean_daily_calories = mean(calories), mean_daily_sleep = mean(total_minutes_asleep))

head(daily_average)
```

-   Determine the user's type based on the category above.

```{r}
user_type <- daily_average %>%
  mutate(user_type = case_when(
    mean_daily_steps < 5000 ~ "sedentary",
    mean_daily_steps >= 5000 & mean_daily_steps < 7499 ~ "lightly active", 
    mean_daily_steps >= 7500 & mean_daily_steps < 9999 ~ "fairly active", 
    mean_daily_steps >= 10000 ~ "very active"
  ))

head(user_type)
```

-   As soon as we have a new column with the user type, we will create a
    data frame with the percentage of each type of user so that the
    distribution can be visualized on a graph.

```{r}
user_type_percent <- user_type %>%
  group_by(user_type) %>%
  summarise(total = n()) %>%
  mutate(totals = sum(total)) %>%
  group_by(user_type) %>%
  summarise(total_percent = total / totals) %>%
  mutate(labels = scales::percent(total_percent))

user_type_percent$user_type <- factor(user_type_percent$user_type , levels = c("very active", "fairly active", "lightly active", "sedentary"))

head(user_type_percent)
```

-   Using the percentage of user types, we create the visualization.

```{r}
user_type_percent %>%
  ggplot(aes(x="",y=total_percent, fill=user_type)) +
  geom_bar(stat = "identity", width = 1)+
  coord_polar("y", start=0)+
  theme_minimal()+
  theme(axis.title.x= element_blank(),
        axis.title.y = element_blank(),
        panel.border = element_blank(), 
        panel.grid = element_blank(), 
        axis.ticks = element_blank(),
        axis.text.x = element_blank(),
        plot.title = element_text(hjust = 0.5, size=14, face = "bold")) +
  scale_fill_manual(values = c("#E76F51","#F4A261", "#2A9D8F", "#264653")) +
  geom_text(aes(label = labels),
            position = position_stack(vjust = 0.5))+
  labs(title="User Type Distribution")
```

#### **2. The Number of Steps Taken and The Time (Minutes) Spent Asleep per Day**

Currently, we're trying to find out what days of the week users are more
active and what days of the week they sleep more. Additionally, we will
verify that the users got the recommended amount of sleep and walked the
recommended number of steps. The weekdays are calculated below based on
our column date. The average steps walked and minutes slept by weekday
are also calculated.

-   The day of the date can be obtained by creating a data
    transformation as below.

```{r}
weekday_steps_sleep <- daily_activity_sleep %>%
  mutate(weekday = weekdays(date))

weekday_steps_sleep$weekday <-ordered(weekday_steps_sleep$weekday, levels=c("Monday", "Tuesday", "Wednesday", "Thursday","Friday", "Saturday", "Sunday"))

weekday_steps_sleep <-weekday_steps_sleep%>%
  group_by(weekday) %>%
  summarize (daily_steps = mean(total_steps), daily_sleep = mean(total_minutes_asleep))

head(weekday_steps_sleep)
```

-   Create the visualization based on the data transformation above.

```{r}
ggarrange(
  ggplot(weekday_steps_sleep) +
    geom_col(aes(weekday, daily_steps), fill = "#E76F51") +
    geom_hline(yintercept = 7500) +
    labs(title = "Daily steps per weekday", x= "", y = "") +
    theme(axis.text.x = element_text(angle = 45,vjust = 0.5, hjust = 1)),
  ggplot(weekday_steps_sleep, aes(weekday, daily_sleep)) +
    geom_col(fill = "#264653") +
    geom_hline(yintercept = 480) +
    labs(title = "Minutes asleep per weekday", x= "", y = "") +
    theme(axis.text.x = element_text(angle = 45,vjust = 0.5, hjust = 1))
)
```

Based on the graphs above, we can draw the following conclusions:

-   Users walk daily the recommended amount of steps of 7500 besides
    Sunday's.

-   Users don't sleep the recommended amount of minutes/ hours - 8
    hours.

#### **3. Hourly Steps, Calories, and Intensities Throughout The Day**

In order to get a better understanding of our analysis, we would like to
know what time of day users are most active. We will use the
`hourly_steps` data frame and separate `date_time` column.

-   Separate date & time from `date_time` column to get hourly data.

```{r}
hourly_cal_intensities_steps <- hourly_cal_intensities_steps %>%
  separate(date_time, into = c("date", "time"), sep= " ") %>%
  mutate(date = ymd(date)) 

head(hourly_cal_intensities_steps)
```

-   Create a visualization of hourly steps.

```{r}
hourly_cal_intensities_steps %>%
  group_by(time) %>%
  summarize(average_steps = mean(step_total)) %>%
  ggplot() +
  geom_col(mapping = aes(x=time, y = average_steps, fill = average_steps)) + 
  labs(title = "Hourly steps throughout the day", x="", y="") + 
  scale_fill_gradient(low = "#2A9D8F", high = "#E76F51")+
  theme(axis.text.x = element_text(angle = 90))
```

-   Create a visualization of hourly calories.

```{r}
hourly_cal_intensities_steps %>%
  group_by(time) %>%
  summarize(average_cals = mean(calories)) %>%
  ggplot() +
  geom_col(mapping = aes(x=time, y = average_cals, fill = average_cals)) + 
  labs(title = "Hourly calories throughout the day", x="", y="") + 
  scale_fill_gradient(low = "#2A9D8F", high = "#E76F51")+
  theme(axis.text.x = element_text(angle = 90))
```

-   Create a visualization of hourly intensities.

```{r}
hourly_cal_intensities_steps %>%
  group_by(time) %>%
  summarize(average_intensity = mean(average_intensity)) %>%
  ggplot() +
  geom_col(mapping = aes(x=time, y = average_intensity, fill = average_intensity)) + 
  labs(title = "Hourly calories throughout the day", x="", y="") + 
  scale_fill_gradient(low = "#2A9D8F", high = "#E76F51")+
  theme(axis.text.x = element_text(angle = 90))
```

#### **4. Correlation Analysis**

We will now determine if there is any correlation between different
variables:

-   Daily steps and daily sleep

```{r}
ggplot(daily_activity_sleep, aes(x=total_steps, y=total_minutes_asleep))+
  geom_point(alpha = 0.3, color = "#264653")+
  geom_smooth(method = "lm", se = TRUE,color = "2A9D8F") + 
  labs(title = "Daily steps vs Minutes asleep", x = "Daily steps", y= "Minutes asleep") +
  theme(panel.background = element_blank(),
        plot.title = element_text( size=14))
```

-   Daily steps and calories

```{r}
ggplot(daily_activity_sleep, aes(x=total_steps, y=calories))+
  geom_point(alpha = 0.3, color = "#264653")+
  geom_smooth(method = "lm", se = TRUE, color = "2A9D8F") + 
  labs(title = "Daily steps vs Calories", x = "Daily steps", y= "Calories") +
  theme(panel.background = element_blank(),
        plot.title = element_text( size=14))
```

#### 5. The Usage of Smart Device

-   **The Number of Days People Use The Device**

    Having seen some trends in activity, sleep, and calories burned, we
    want to find out how often the users in our sample use their
    devices. We can then figure out what features are beneficial to
    smart devices and plan our marketing strategy.

    Assuming that the date interval is 31 days, we will calculate the
    number of users who use their smart devices every day, by
    classifying our sample into three groups:

    -   High use: Users who use their device between 21 and 31 days.

    -   Moderate use: Users who use their device between 10 and 20 days.

    -   Low use: Users who use their device between 1 and 10 days.

    First, we will create a new data frame grouping by `id`, calculating
    number of days used and creating a new column with the
    classification explained above.

    ```{r}
    daily_use <- daily_activity_sleep %>%
      filter(total_steps > 200) %>% 
      group_by(id) %>%
      summarize(days_used=sum(n())) %>%
      mutate(usage = case_when(
        days_used >= 1 & days_used <= 10 ~ "Low Use",
        days_used >= 11 & days_used <= 20 ~ "Moderate Use", 
        days_used >= 21 & days_used <= 31 ~ "High Use", 
      )) %>% 
      mutate(usage = factor(usage, level = c('Low Use','Moderate Use','High Use'))) %>% 
      group_by(usage) %>% 
      arrange(perc) %>% 
      mutate(perc = scales::percent(perc))

    head(daily_use)
    ```

    ```{r}
    daily_use <- daily_activity %>% 
      inner_join(daily_use, by = "id") %>%
      group_by(usage) %>% 
      summarise(participants = n_distinct(id)) %>% 
      mutate(perc = participants/sum(participants)) %>% 

    head(daily_use)
    ```

-   In order to better visualize the results in the graph, we will now
    create a percentage data frame. The usage levels will also be
    ordered.

    ```{r}
    daily_use_percent <- daily_use %>%
      group_by(usage) %>%
      summarise(total = n()) %>%
      mutate(totals = sum(total)) %>%
      group_by(usage) %>%
      summarise(total_percent = total / totals) %>%
      mutate(labels = scales::percent(total_percent))

    daily_use_percent$usage <- factor(daily_use_percent$usage, levels = c("high use", "moderate use", "low use"))

    head(daily_use_percent)
    ```

    Once we have our new table, we can create our plot:

    ```{r}
    daily_use_percent %>%
      ggplot(aes(x="",y=total_percent, fill=usage)) +
      geom_bar(stat = "identity", width = 1)+
      coord_polar("y", start=0)+
      theme_minimal()+
      theme(axis.title.x= element_blank(),
            axis.title.y = element_blank(),
            panel.border = element_blank(), 
            panel.grid = element_blank(), 
            axis.ticks = element_blank(),
            axis.text.x = element_blank(),
            plot.title = element_text(hjust = 0.5, size=14, face = "bold")) +
      geom_text(aes(label = labels),
                position = position_stack(vjust = 0.5))+
      scale_fill_manual(values = c("#E76F51","#F4A261","#2A9D8F"),
                        labels = c("High use - 21 to 31 days",
                                   "Moderate use - 11 to 20 days",
                                   "Low use - 1 to 10 days"))+
      labs(title="Daily use of smart device")
    ```

### **Act Phase (Conclusion)**

Bellabeat's mission is to empower women by providing them with the data
to discover themselves.

**1) *"More than just a sports tracker"***

Users need to understand that to get more out of the device, frequent
use would facilitate better data collection that can help bring more
insights and actionable recommendations. More research must be done to
understand why they 'Lower' and 'Moderate Use' groups are not wearing
their devices more often. One such conclusion could point at the users
perception that the devices are only useful when worn during excercise
or physical activities.

Marketing could focus on how the coupling of the 'wellness' and 'sports'
elements can help bring a more holistic understanding of their wellness
and lifestyle habits.

Product designers could focus on gamifying features to encourage more
frequent use in the devices. Having smoother app to devices integration
and syncing will help facilitate this. Like how GoPro differentiates
themselves from the rest of the pack by having a comprehensive and user
friendly app design.

**2) Socialization Features**

With data obtained from hourly and daily metrics. Users can be engaged
further by receiving prompts on how they are doing at the point of the
day or week when compared to others in their age/location/gender group.
Congratulating users on meeting their targets or encouraging them to go
the extra mile. These could be some of the prompts/push-notifications:

-   Good job! You've taken as many steps as those in your age group!

-   Keep it up! You're almost as active as you were this time last week!

-   You're not alone in this! You're in the 50th percentile of most
    active minutes this week!

-   Rest up! A good recovery is just as important as a good run.

**3) Expanding device line-up**

Most of Bellabeat's products are devices operate discretely without a
screen. Most information have to be viewed directly from the app and not
from the devices itself. A whopping 94% of users (Moderate & High Users)
wear their devices more than 15 days in a month. They would wear their
devices throughout the day as well as their during their workouts. This
presents an opportunity for Bellabeat to introduce a new line of
smartwatches with touch screens focused on 'every day' wear. Together
with their focus on women-centric products, this choice seems obvious in
differentiating them from the rest of the pack.

**4) More inclusive studies needed**

Insights on women's health and wellness were not represented in this
study. Data on how women use these devices and compared to men can
reveal crucial insights for Bellabeat, especially on features such as
menstrual health. I alone, a cisgender man, would bring a limited if not
harmful assessment of this data.
